Renormalization

### 2014-05-20 papers/renormalization/paper.tex
Small but important change to RG paper Tested on elliot. Took 0:19.43.
clean up summation notation and write about multiplicity Tested on bennet. Took 0:16.41.

  Z &= \sum_i^{\scriptscriptstyle\begin{array}{c}\text{\tiny all}\\\text{\tiny microstates}\end{array}}

@@

@@

  Z_j &= \sum_i^{\scriptscriptstyle\begin{array}{c}
                  \text{\tiny microstates}\\
                  \text{\tiny periodic }2^j\lambda\\
                 \end{array}}

@@

@@

  Z_j(N) &= \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
                  \text{\tiny occupancies}\\
                  \text{\tiny of octants}\\
                 \end{array}}
       \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
                  \text{\tiny microstates}\\
                  \text{\tiny within octants}\\
                 \end{array}}

@@

    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
                  \text{\tiny occupancies}\\
                  \text{\tiny of octants}\\
                 \end{array}}

@@

       \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
                  \text{\tiny microstates}\\
                  \text{\tiny within octants}\\
                 \end{array}}

@@

    \sum_{AAN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
                  \text{\tiny occupancies}\\
                  \text{\tiny of octants}\\
                 \end{array}}

@@

    \sum_{NN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
                  \text{\tiny occupancies}\\
                  \text{\tiny of octants}\\
                 \end{array}}

@@

    \sum_{BN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
                  \text{\tiny occupancies}\\
                  \text{\tiny of octants}\\
                 \end{array}}

@@

The flavor of the approximation used by White and others in GRG is now
clarified.  We begin by assuming that we already have an adequate (and
tractable) approximation for $F_{j-1}$.  Thus the denominator
partition function is readily computable.  We further simplify the
numerator by multiplying and dividing by the same quantity, and
simplify the notation by defining
\begin{align}
  F_j(N_1\cdots N_8) &= \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
                  \text{\tiny microstates}\\
                  \text{\tiny within octants}\\
                 \end{array}}
       e^{-\beta E_{i_1\cdots i_8}^{(j)}}
\end{align}
and thus rewrite the energy difference as
\begin{multline}
  F_j(n) - F_{j-1}(n) =\\
  -kT\log\left(\frac{
    \sum_{CN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
                  \text{\tiny occupancies}\\
                  \text{\tiny of octants}\\
                 \end{array}}
       g_{N_1\cdots N_8}
       e^{-\beta \big(F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) +
           \big(F_j(N_1\cdots N_8) - F_{j-1}(N_1) - \cdots - F_{j-1}(N_8)\big) \big)}
  }{
    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
                  \text{\tiny occupancies}\\
                  \text{\tiny of octants}\\
                 \end{array}}
       g_{N_1\cdots N_8}
       e^{-\beta \left( F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) \right)}
  }
  \right)
\end{multline}
At this point we can recognize and identify two energy differences as
\begin{align}
  \bar{F}(N_1\cdots N_8) &= F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) \\
  \bar{U}(N_1\cdots N_8) &= F_j(N_1\cdots N_8)
  - \left(F_{j-1}(N_1) + \cdots + F_{j-1}(N_8) \right)
\end{align}
and using these definitions can rewrite our energy difference in a
more compact form:
\begin{multline}
  F_j(n) - F_{j-1}(n) =\\
  -kT\log\left(\frac{
    \sum_{CN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
                  \text{\tiny occupancies}\\
                  \text{\tiny of octants}\\
                 \end{array}}
       g_{N_1\cdots N_8}
       e^{-\beta \big( \bar{F}(N_1\cdots N_8) + \bar{U}(N_1\cdots N_8)
         \big)}
  }{
    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
                  \text{\tiny occupancies}\\
                  \text{\tiny of octants}\\
                 \end{array}}
       g_{N_1\cdots N_8}
       e^{-\beta \bar{F}(N_1\cdots N_8)}
  }
  \right)
\end{multline}
At this point, we can recognize $\bar{F}(N_1\cdots N_8)$ as the free
energy change due to oscillations \emph{if we neglect the actual
  energy of the oscillations themselves, but instead make a ``local''
  approximation}.  Meanwhile $\bar{U}(N_1\cdots N_8)$ is the free
energy change due to the energy of oscillations.

\fixme{I have now ``defined'' the terms used by Forte in his
  notation of GRG.  On the to-do list: }
\begin{enumerate}
\item \fixme{Establish what $g_{N_1\cdots N_8}$ is.}
\item \fixme{Elaborate on the approximation (ignoring corners) that
  leads from a $\sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
    \text{\tiny occupancies}\\ \text{\tiny of octants}\\
                 \end{array}}
       g_{N_1\cdots N_8}$ to an integral over density
       amplitudes.}
\item \fixme{Develop an approach for approximating $\bar{U}$.  Perhaps
  is $F_{j-1}$ is expressed as a functional, we can use a non-local
  DFT to approximate $\bar U_j$?  The standard GRG approach seems to
  be to use a simple mean-field or thermodynamic perturbation theory
  approach to approximate $\bar U$, suggesting that a simple
  (standard) DFT might work just fine for this.}
\end{enumerate}

>>>>

  Z &= \sum_i^{\text{\tiny\begin{minipage}{5em}
      \begin{center}
        \tiny all microstates\end{center}\end{minipage}}}

@@

\newcommand\sumstatesperiodic[1]{\sum_i^{\text{\tiny\begin{minipage}{6em}
      \begin{center}
        \tiny microstates periodic $#1$\end{center}\end{minipage}}}}

@@

  Z_j &= \sumstatesperiodic{2^j\lambda}

@@

\newcommand\sumoccupanciesofoctants{
  \sum_{N_1\cdots N_8}^{\text{\tiny\begin{minipage}{5em}
      \begin{center}
      \tiny numbers in octants\end{center}\end{minipage}}}
\newcommand\summicrostatesofoctants{
  \sum_{i_1\cdots i_8}^{\text{\tiny\begin{minipage}{4em}
      \begin{center}
      \tiny microstates\\within octants\end{center}\end{minipage}}}

@@

  Z_j(N) &= \sumoccupanciesofoctants g_{N_1\cdots N_8}
       \summicrostatesofoctants

@@

    \sumoccupanciesofoctants

@@

       \summicrostatesofoctants

@@

    \sumoccupanciesofoctants

@@

    \sumoccupanciesofoctants

@@

    \sumoccupanciesofoctants

@@

The flavor of the approximation used by White and others in GRG is now
clarified.  We begin by assuming that we already have an adequate (and
tractable) approximation for $F_{j-1}$.  Thus the denominator
partition function is readily computable.  We further simplify the
numerator by multiplying and dividing by the same quantity, and
simplify the notation by defining
\begin{align}
  F_j(N_1\cdots N_8) &= \summicrostatesofoctants
       e^{-\beta E_{i_1\cdots i_8}^{(j)}}
\end{align}
and thus rewrite the energy difference as
\begin{multline}
  F_j(n) - F_{j-1}(n) =\\
  -kT\log\left(\frac{
    \sumoccupanciesofoctants
       g_{N_1\cdots N_8}
       e^{-\beta \big(F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) +
           \big(F_j(N_1\cdots N_8) - F_{j-1}(N_1) - \cdots - F_{j-1}(N_8)\big) \big)}
  }{
    \sumoccupanciesofoctants
       g_{N_1\cdots N_8}
       e^{-\beta \left( F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) \right)}
  }
  \right)
\end{multline}
At this point we can recognize and identify two energy differences as
\begin{align}
  \bar{F}(N_1\cdots N_8) &= F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) \\
  \bar{U}(N_1\cdots N_8) &= F_j(N_1\cdots N_8)
  - \left(F_{j-1}(N_1) + \cdots + F_{j-1}(N_8) \right)
\end{align}
and using these definitions can rewrite our energy difference in a
more compact form:
\begin{multline}
  F_j(n) - F_{j-1}(n) =\\
  -kT\log\left(\frac{
    \sumoccupanciesofoctants
       g_{N_1\cdots N_8}
       e^{-\beta \big( \bar{F}(N_1\cdots N_8) + \bar{U}(N_1\cdots N_8)
         \big)}
  }{
    \sumoccupanciesofoctants
       g_{N_1\cdots N_8}
       e^{-\beta \bar{F}(N_1\cdots N_8)}
  }
  \right)
\end{multline}
At this point, we can recognize $\bar{F}(N_1\cdots N_8)$ as the free
energy change due to oscillations \emph{if we neglect the actual
  energy of the oscillations themselves, but instead make a ``local''
  approximation}.  Meanwhile $\bar{U}(N_1\cdots N_8)$ is the free
energy change due to the energy of oscillations.

\fixme{I have now ``defined'' the terms used by Forte in his
  notation of GRG.  On the to-do list: }
\begin{enumerate}
\item \fixme{Establish what $g_{N_1\cdots N_8}$ is.}
  \begin{align}
    g_{N_1\cdots N_8} &=
    {N \choose N_1} {N - N_1 \choose N_2} {N - N_1 - N_2 \choose N_3}
    \cdots \\
    &\propto \sum_{i=1}^{8} N_i!
    \\
    &\approx \sum_{i=1}^{8} (N_i-1) e^{N_i}
  \end{align}
  This combinatoric term ends up showing up as the ideal gas free
  energy term.
\item \fixme{Elaborate on the approximation (ignoring corners) that
  leads from a $\sumoccupanciesofoctants g_{N_1\cdots N_8}$ to an
  integral over density amplitudes.}
\item \fixme{Develop an approach for approximating $\bar{U}$.  Perhaps
  is $F_{j-1}$ is expressed as a functional, we can use a non-local
  DFT to approximate $\bar U_j$?  The standard GRG approach seems to
  be to use a simple mean-field or thermodynamic perturbation theory
  approach to approximate $\bar U$, suggesting that a simple
  (standard) DFT might work just fine for this.}
\end{enumerate}
